Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-02-21 LaTeX/Turnover_bioRXiv.tex
tex, sup figs pdf, and mathematica pdf for bioRxiv submission

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}
%% Include all macros below

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Model Outline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
%\centering
%\centerline{\includegraphics[width=\linewidth]{Sex_determination_outline.pdf}}
\caption{
{\bf Outline of model features.}
Panel A: Recombination rate parameters between the ancestral-sex-determining locus ($\mathbf{X}$, here assumed to have alleles X and Y), a locus under selection ($\mathbf{A}$, with alleles $A$ and $a$), and a new sex-determining locus ($\mathbf{M}$, with alleles $M$ and $m$). 
If $r<1/2$, then associations between ancestral sex-determining alleles and selected alleles can be maintained past recombination in males. 
Panel B: Haploid selection is often sex-specific, occurring during haploid production or competition in either males or females. 
For example, haploid selection in males only is represented by the dashed circle.  
If X or Y alleles remain associated with alleles that experience haploid selection in males ($r<1/2$), then zygotic sex ratios can become biased because either X- or Y-bearing male gametes/gametophytes will be more abundant after haploid selection. 
However, the zygotic sex ratio is not biased by male haploid selection in ZW sex-determining systems. 
Similarly, zygotic sex ratio biases can occur if haploid selected alleles are associated with new sex-determining alleles ($R<1/2$). 
Panel C: During cis-GSD transitions (XY to XY or ZW to ZW), a neo-Y allele spreads to pseudo-fixation (its maximum frequency among male gametes) and the ancestral Y allele is lost. 
Panel D: During trans-GSD transitions (XY to ZW or ZW to XY), a neo-W allele spreads to pseudo-fixation (its maximum frequency among female gametes) and the ancestral X allele is lost. 
Neo-W alleles allow Y-associated alleles into females, which may impede or aid their spread. 
}
\label{fig:model_outline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[!ht]
\centering
\smallskip
%\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W alleles}
\caption{ {\bf Parameters determining invasion of mutant neo-Y and neo-W alleles into an ancestrally XY system}}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   $m$ is a neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\Lambda_{Y'A}^{(XY)} = {\left( 2 \zeta \right)}^{-1} \left[\hat{p}_X^\female w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male} + (1-\hat{p}_X^\female) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} (1+\alpha_{\Delta}^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right) $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\Lambda_{Y'a}^{(XY)} = {\left( 2 \zeta \right)}^{-1} \left[(1-\hat{p}_X^\female) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male} + \hat{p}_X^\female w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1 - \alpha_{\Delta}^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{Y'A}^{(XY)} = R {\left( 2 \zeta \right)}^{-1} \left[ (1-\hat{p}_X^\female) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} (1+\alpha_{\Delta}^{\male}) \right]/  \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right)   $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{Y'a}^{(XY)} = R {\left( 2 \zeta \right)}^{-1} \left[   \hat{p}_X^\female w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} (1 - \alpha_\Delta^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right)  $\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  $m$ is a neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\Lambda_{W'A}^{(XY)} = {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ \bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}(1+\alpha_{\Delta}^{\female})\right]/ \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\Lambda_{W'a}^{(XY)} = {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ (1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha_{\Delta}^{\female})\right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{W'A}^{(XY)} = R {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ (1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} (1+\alpha_{\Delta}^{\female}) \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right)  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{W'a}^{(XY)} = R {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ \bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} (1-\alpha_{\Delta}^{\female}) \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [1ex]
  \hline \hline 
   \end{tabular}
   \begin{flushleft} 
$\bar{p}^{\male}=(1-\hat{q})\hat{p}_X^\male + \hat{q}\hat{p}_Y^\male$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes.
%      \item $R$ is the probability of recombination between loci $\mathbf{A}$ and $\mathbf{M}$.
$\zeta$ is the zygotic sex ratio (fraction male).
$\bar{w}^{\Hermaphrodite}_{D}$ is the mean fitness of diploids of sex $\Hermaphrodite$.
$\bar{w}_H^\Hermaphrodite$ is the mean fitness of haploids from sex $\Hermaphrodite$, see \nameref{tab:meanfitnesses}.
\end{flushleft}
  \label{tab:haplotype_growth}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Tighter linkage allows looser sex-linkage to evolve, even in absence of haploid selection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\centerline{\includegraphics[width=1.5\linewidth]{PositionPlot_SexAntagTighter}}
\caption{
{\bf Transitions between XY and ZW systems can occur even when the new sex-determining locus is less tightly linked to a locus under sexually-antagonistic selection (even without haploid selection).}
In panel A, linkage is loose enough relative to selection that the analytical results assuming weak selection hold, and a neo-W allele can only invade when it arises at a locus more tightly linked with the selected locus ($R<r$; shaded region).
In panel B, linkage is tight enough relative to selection that the analytical results assuming weak selection do not hold, and a neo-W can invade even when it is less tightly linked with the selected locus ($r<R$; unshaded region around \text{*}).
In panel C we vary the recombination rate between the neo-W and the selected locus ($R$) for a fixed recombination rate between the ancestral sex-determining locus and the selected locus ($r=0.005$). 
Coloured markers show recombination rates for which the temporal dynamics of invasion are plotted in the inset, demonstrating that neo-W alleles can reach pseudo-fixation if they are more (black) or less (red) closely linked to a locus experiencing sexually-antagonistic selection. 
A very loosely linked neo-W does not spread in this case (blue and green lines overlap and go to 0 in inset). 
Fitness parameters are: $w_{AA}^\female = 1.05$, $w_{aa}^\male = 1.2$, $w_{aa}^\female = w_{AA}^\male = 0.85$, $w_{Aa}^\Hermaphrodite = 1$,  $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
%\textcolor{red}{consider removing panel A, which is repeated in Fig 3.}
%\textcolor{blue}{probably better to remove from panel 3, eh?}
}
\label{fig:SexAntagTighter}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Both neo-W haplotypes can spread, in the absence of haploid selection, when there is overdominance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\centerline{\includegraphics[width=1.25\linewidth]{Region_plot_combined_Mike}}
%\centerline{\includegraphics[width=1.25\linewidth]{Region_plot_combined}}
\caption{
{\bf When the ancestral XY locus is tightly linked to a locus under selection ($r=0$), one or both neo-W haplotypes can spread. }
We vary the fitness of male homozygotes relative to heterozygotes ($w_{Aa}^\Hermaphrodite=1$) and only consider stable equilibria at which both $\mathbf{A}$ locus alleles are maintained and the $a$ allele is initially fixed on Y backgrounds, region outlined. 
Here, selection in females can favour the $A$ allele (panel A, $w_{aa}^\female=0.85$, $w_{AA}^\female=1.05$), favour the $a$ allele (panel B, $w_{aa}^\female=1.05$, $w_{AA}^\female=0.85$), or be overdominant (panel C, $w_{aa}^\female=w_{AA}^\female=0.6$). 
If either haplotypic growth rate ($\Lambda_{W'A}^{(XY)}$ or $\Lambda_{W'a}^{(XY)}$) is greater than one, then a rare neo-W allele can spread for, at least, some values of $R>r$. 
The parameter values marked with an asterisk correspond to the fitnesses used in Fig \ref{fig:SexAntagTighter}C. 
Where both haplotypic growth rates are greater than one, a neo-W will spread when rare, regardless of linkage with the selected locus (for any $R$). 
\nameref{fig:positionOverdominance} shows the dynamics arising with the parameters marked with a dagger. 
Here, there is no haploid selection $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
}
\label{fig:regionplots}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%A neo-W can invade an XY system under a large number of selective regimes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\includegraphics[width=\linewidth]{PositionPlot}
\caption{
{\bf Ploidally-antagonistic selection allows a less tightly linked neo-W allele to invade.}
In panel A, male drive ($\alpha^\male_{\Delta} = -1/20$, $t^\Hermaphrodite = \alpha^\female_{\Delta} = 0$) opposes selection in diploids (no sex-differences: $s^\Hermaphrodite = 1/10$, $h^\Hermaphrodite = 7/10$), in which case the new sex-determining allele can invade regardless of its linkage with the selected locus ($R$).   
In panel B, gametic competition in males ($t^\male = -1/10$, $t^\female = \alpha^\Hermaphrodite_{\Delta} = 0$) opposes selection in diploids (sex-differences: $s^\male = 3/20$, $s^\female = 1/20$, $h^\Hermaphrodite = 7/10$), in which case the new sex-determining allele can once again invade regardless of $R$.
}
\label{fig:Combination_Centimorgans}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[!ht]
\centering
\smallskip
\caption{Invasion conditions for a neo-W allele at an unlinked locus ($R=1/2$) into an ancestral XY system with linkage ($r<1/2$) and one form of haploid selection}
\begin{tabular}{l l c }
\hline\hline
Scenario &  Assumptions & neo-W spreads ($\lambda_{W'}^{(XY)}>1$) if \\ [0.5ex] \hline
%  $r=1/2$, $R<1/2$ & Always \\ [0.5ex] \hline
%  $r<1/2$, $R=1/2$ &  \\ [0.5ex] \hline
\noalign{\vskip 1mm}
  male drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\female_{\Delta}=0$ & $s^\female s^\male>0$ \\ [0.5ex]
 female drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\male_{\Delta}=0$ & $s^\female s^\male>0$ \\ [0.5ex]
 sperm competition only &  $h^\male=h^\female$, $t^\female=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\female(s^\male-s^\female)>0$ \\ [0.5ex]
  egg competition only & $h^\male=h^\female$, $t^\male=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\male(s^\female-s^\male)>0$ \\ [0.5ex]
  \hline \hline
  \label{tab:specialcases}
 \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Sex ratio selection is not a good predictor of neo-W invasion into an XY system
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\includegraphics[width=\linewidth]{Temporal_SR}
\caption{
{\bf Fisherian sex ratio selection alone is not a good predictor of turnover between sex-determining systems.}
In this figure, selection is ploidally antagonistic with haploid selection favouring the $a$ allele during male meiosis.
In panel A, male meiotic drive in an ancestral XY system causes a male bias (see Fig \ref{fig:model_outline}B), allowing a neo-W to invade and replace the ancestral sex-determining system (inset shows neo-W frequency among female gametes, reaching pseudo-fixation), which balances the zygotic sex ratio.
In panel B, male drive in an ancestral ZW system has no effect on the zygotic sex ratio yet a neo-Y can invade and replace the ancestral sex-determining system (inset shows neo-Y frequency among male gametes, reaching pseudo-fixation). 
%When $R<1/2$, the neo-Y becomes associated with the allele favoured by drive, causing the zygotic sex ratio to become biased, hence the frequency of the neo-Y at pseudo-fixation can be higher than $0.5$ (inset).
Parameters:  $s^\female =s^\male = 0.2$, $h^\female = h^\male = 0.7$, $t^\female = t^\male = \alpha^\female_\Delta = 0$, $\alpha^\male_\Delta = -0.1$, $r=0.02$.
}
\label{fig:SexRatioBad}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph*{S2 Table}
\label{tab:meanfitnesses}
{\bf Mean fitnesses and zygotic sex ratio in the resident population ($M$ fixed, XY sex determination). }

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,footskip=0.75in]{geometry}

@@

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 6.25in 
\textheight 8.75in

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
%\fancyheadoffset[L]{2.25in}
%\fancyfootoffset[L]{2.25in}
%\lfoot{\sf PLOS}
%% Include all macros below

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Model Outline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{../Plots/Sex_determination_outline.pdf}}
\caption{
{\bf Outline of model features.}
Panel A: Recombination rate parameters between the ancestral-sex-determining locus ($\mathbf{X}$, here assumed to have alleles X and Y), a locus under selection ($\mathbf{A}$, with alleles $A$ and $a$), and a new sex-determining locus ($\mathbf{M}$, with alleles $M$ and $m$). 
If $r<1/2$, then associations between ancestral sex-determining alleles and selected alleles can be maintained past recombination in males. 
Panel B: Haploid selection is often sex-specific, occurring during haploid production or competition in either males or females. 
For example, haploid selection in males only is represented by the dashed circle.  
If X or Y alleles remain associated with alleles that experience haploid selection in males ($r<1/2$), then zygotic sex ratios can become biased because either X- or Y-bearing male gametes/gametophytes will be more abundant after haploid selection. 
However, the zygotic sex ratio is not biased by male haploid selection in ZW sex-determining systems. 
Similarly, zygotic sex ratio biases can occur if haploid selected alleles are associated with new sex-determining alleles ($R<1/2$). 
Panel C: During cis-GSD transitions (XY to XY or ZW to ZW), a neo-Y allele spreads to pseudo-fixation (its maximum frequency among male gametes) and the ancestral Y allele is lost. 
Panel D: During trans-GSD transitions (XY to ZW or ZW to XY), a neo-W allele spreads to pseudo-fixation (its maximum frequency among female gametes) and the ancestral X allele is lost. 
Neo-W alleles allow Y-associated alleles into females, which may impede or aid their spread. 
}
\label{fig:model_outline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[!ht]
%\centering
\smallskip
%\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W alleles}
\caption{ {\bf Parameters determining invasion of mutant neo-Y and neo-W alleles into an ancestrally XY system}}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   $m$ is a neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\Lambda_{Y'A}^{(XY)} = {\left( 2 \zeta \right)}^{-1} \left[\hat{p}_X^\female w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male} + (1-\hat{p}_X^\female) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} (1+\alpha_{\Delta}^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right) $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\Lambda_{Y'a}^{(XY)} = {\left( 2 \zeta \right)}^{-1} \left[(1-\hat{p}_X^\female) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male} + \hat{p}_X^\female w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1 - \alpha_{\Delta}^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{Y'A}^{(XY)} = R {\left( 2 \zeta \right)}^{-1} \left[ (1-\hat{p}_X^\female) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} (1+\alpha_{\Delta}^{\male}) \right]/  \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right)   $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{Y'a}^{(XY)} = R {\left( 2 \zeta \right)}^{-1} \left[   \hat{p}_X^\female w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} (1 - \alpha_\Delta^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\male}_{D} \right)  $\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  $m$ is a neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\Lambda_{W'A}^{(XY)} = {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ \bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}(1+\alpha_{\Delta}^{\female})\right]/ \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\Lambda_{W'a}^{(XY)} = {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ (1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha_{\Delta}^{\female})\right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{W'A}^{(XY)} = R {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ (1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} (1+\alpha_{\Delta}^{\female}) \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right)  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\chi_{W'a}^{(XY)} = R {\left[ 2 (1 - \zeta) \right]}^{-1} \left[ \bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} (1-\alpha_{\Delta}^{\female}) \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^{\female}_{D} \right) $ \\ [1ex]
  \hline \hline 
   \end{tabular}
   \begin{flushleft} 
$\bar{p}^{\male}=(1-\hat{q})\hat{p}_X^\male + \hat{q}\hat{p}_Y^\male$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes.
%      \item $R$ is the probability of recombination between loci $\mathbf{A}$ and $\mathbf{M}$.
$\zeta$ is the zygotic sex ratio (fraction male).
$\bar{w}^{\Hermaphrodite}_{D}$ is the mean fitness of diploids of sex $\Hermaphrodite$.
$\bar{w}_H^\Hermaphrodite$ is the mean fitness of haploids from sex $\Hermaphrodite$, see \nameref{tab:meanfitnesses}.
\end{flushleft}
  \label{tab:haplotype_growth}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Tighter linkage allows looser sex-linkage to evolve, even in absence of haploid selection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.25\linewidth]{../Plots/PositionPlot_SexAntagTighter}}
\caption{
{\bf Transitions between XY and ZW systems can occur even when the new sex-determining locus is less tightly linked to a locus under sexually-antagonistic selection (even without haploid selection).}
In panel A, linkage is loose enough relative to selection that the analytical results assuming weak selection hold, and a neo-W allele can only invade when it arises at a locus more tightly linked with the selected locus ($R<r$; shaded region).
In panel B, linkage is tight enough relative to selection that the analytical results assuming weak selection do not hold, and a neo-W can invade even when it is less tightly linked with the selected locus ($r<R$; unshaded region around \text{*}).
In panel C we vary the recombination rate between the neo-W and the selected locus ($R$) for a fixed recombination rate between the ancestral sex-determining locus and the selected locus ($r=0.005$). 
Coloured markers show recombination rates for which the temporal dynamics of invasion are plotted in the inset, demonstrating that neo-W alleles can reach pseudo-fixation if they are more (black) or less (red) closely linked to a locus experiencing sexually-antagonistic selection. 
A very loosely linked neo-W does not spread in this case (blue and green lines overlap and go to 0 in inset). 
Fitness parameters are: $w_{AA}^\female = 1.05$, $w_{aa}^\male = 1.2$, $w_{aa}^\female = w_{AA}^\male = 0.85$, $w_{Aa}^\Hermaphrodite = 1$,  $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
%\textcolor{red}{consider removing panel A, which is repeated in Fig 3.}
%\textcolor{blue}{probably better to remove from panel 3, eh?}
}
\label{fig:SexAntagTighter}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Both neo-W haplotypes can spread, in the absence of haploid selection, when there is overdominance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\centerline{\includegraphics[width=1.25\linewidth]{Region_plot_combined_Mike}}
\centerline{\includegraphics[width=1.25\linewidth]{../Plots/Region_plot_combined}}
\caption{
{\bf When the ancestral XY locus is tightly linked to a locus under selection ($r=0$), one or both neo-W haplotypes can spread. }
We vary the fitness of male homozygotes relative to heterozygotes ($w_{Aa}^\Hermaphrodite=1$) and only consider stable equilibria at which both $\mathbf{A}$ locus alleles are maintained and the $a$ allele is initially fixed on Y backgrounds, region outlined. 
Here, selection in females can favour the $A$ allele (panel A, $w_{aa}^\female=0.85$, $w_{AA}^\female=1.05$), favour the $a$ allele (panel B, $w_{aa}^\female=1.05$, $w_{AA}^\female=0.85$), or be overdominant (panel C, $w_{aa}^\female=w_{AA}^\female=0.6$). 
If either haplotypic growth rate ($\Lambda_{W'A}^{(XY)}$ or $\Lambda_{W'a}^{(XY)}$) is greater than one, then a rare neo-W allele can spread for, at least, some values of $R>r$. 
The parameter values marked with an asterisk correspond to the fitnesses used in Fig \ref{fig:SexAntagTighter}C. 
Where both haplotypic growth rates are greater than one, a neo-W will spread when rare, regardless of linkage with the selected locus (for any $R$). 
\nameref{fig:positionOverdominance} shows the dynamics arising with the parameters marked with a dagger. 
Here, there is no haploid selection $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
}
\label{fig:regionplots}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%A neo-W can invade an XY system under a large number of selective regimes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{../Plots/PositionPlot}
\caption{
{\bf Ploidally-antagonistic selection allows a less tightly linked neo-W allele to invade.}
In panel A, male drive ($\alpha^\male_{\Delta} = -1/20$, $t^\Hermaphrodite = \alpha^\female_{\Delta} = 0$) opposes selection in diploids (no sex-differences: $s^\Hermaphrodite = 1/10$, $h^\Hermaphrodite = 7/10$), in which case the new sex-determining allele can invade regardless of its linkage with the selected locus ($R$).   
In panel B, gametic competition in males ($t^\male = -1/10$, $t^\female = \alpha^\Hermaphrodite_{\Delta} = 0$) opposes selection in diploids (sex-differences: $s^\male = 3/20$, $s^\female = 1/20$, $h^\Hermaphrodite = 7/10$), in which case the new sex-determining allele can once again invade regardless of $R$.
}
\label{fig:Combination_Centimorgans}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[!ht]
%\centering
\smallskip
\caption{Invasion conditions for a neo-W allele at an unlinked locus ($R=1/2$) into an ancestral XY system with linkage ($r<1/2$) and one form of haploid selection}
\begin{tabular}{l l c }
\hline\hline
Scenario &  Assumptions & neo-W spreads ($\lambda_{W'}^{(XY)}>1$) if \\ [0.5ex] \hline
%  $r=1/2$, $R<1/2$ & Always \\ [0.5ex] \hline
%  $r<1/2$, $R=1/2$ &  \\ [0.5ex] \hline
\noalign{\vskip 1mm}
  male drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\female_{\Delta}=0$ & $s^\female s^\male>0$ \\ [0.5ex]
 female drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\male_{\Delta}=0$ & $s^\female s^\male>0$ \\ [0.5ex]
 sperm competition only &  $h^\male=h^\female$, $t^\female=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\female(s^\male-s^\female)>0$ \\ [0.5ex]
  egg competition only & $h^\male=h^\female$, $t^\male=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\male(s^\female-s^\male)>0$ \\ [0.5ex]
  \hline \hline
  \label{tab:specialcases}
 \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Sex ratio selection is not a good predictor of neo-W invasion into an XY system
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{../Plots/Temporal_SR}
\caption{
{\bf Fisherian sex ratio selection alone is not a good predictor of turnover between sex-determining systems.}
In this figure, selection is ploidally antagonistic with haploid selection favouring the $a$ allele during male meiosis.
In panel A, male meiotic drive in an ancestral XY system causes a male bias (see Fig \ref{fig:model_outline}B), allowing a neo-W to invade and replace the ancestral sex-determining system (inset shows neo-W frequency among female gametes, reaching pseudo-fixation), which balances the zygotic sex ratio.
In panel B, male drive in an ancestral ZW system has no effect on the zygotic sex ratio yet a neo-Y can invade and replace the ancestral sex-determining system (inset shows neo-Y frequency among male gametes, reaching pseudo-fixation). 
%When $R<1/2$, the neo-Y becomes associated with the allele favoured by drive, causing the zygotic sex ratio to become biased, hence the frequency of the neo-Y at pseudo-fixation can be higher than $0.5$ (inset).
Parameters:  $s^\female =s^\male = 0.2$, $h^\female = h^\male = 0.7$, $t^\female = t^\male = \alpha^\female_\Delta = 0$, $\alpha^\male_\Delta = -0.1$, $r=0.02$.
}
\label{fig:SexRatioBad}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph*{S2 Table}
\label{tab:meanfitnesses}
{\bf Mean fitnesses and zygotic sex ratio in the resident population ($M$ fixed, XY sex determination).}
